Loading and preprocessing the data
-----------------------------------

```{r}
library(plyr) 
library(ggplot2) 
activity <- read.csv("activity.csv") # read in the data 
activity$date <- as.POSIXct(activity$date) # set the dates to POSIXct
```
What is mean total number of steps taken per day?
-------------------------------------------------

## Calculate the total number of steps taken per day
```{r}
dailysteps <- aggregate(activity$steps, by = list(activity$date), sum, na.rm=TRUE)
names(dailysteps) <- c("Date", "steps")
```

## Make a histogram of the total number of steps taken each day
```{r}
qplot(steps, data = dailysteps, geom="histogram", xlab = "Daily Number of Steps", binwidth = 300)
```

```{r}
mean.steps <- mean(dailysteps$steps) 
median.steps <- median(dailysteps$steps)
```
The mean number of steps each day is 9354 
The median number of steps each day is 10395

What is the average daily activity pattern?
-------------------------------------------

## df of the mean and median number of steps taken, averaged across all days (y-axis)

```{r}
intsteps <- aggregate(activity$steps, by = list(activity$interval), mean, na.rm=TRUE)
intstepsmed <- aggregate(activity$steps, by = list(activity$interval), median, na.rm=TRUE)
intsteps <- cbind(intsteps[], intstepsmed$x)
```
## Tidy the df names and round the numbers

```{r}
names(intsteps) = c("interval","mean.steps", "median.steps")
intsteps$mean.steps <- round(intsteps$mean.steps)
intsteps$median.steps <- round(intsteps$median.steps)
ggplot(intsteps, aes(x = interval, y = mean.steps)) + geom_line()
```
As can be seen from the plots above the user appears to start walking around later at weekends (maybe they sleep later). Once they are awake they tend to take more steps and to be walking around (awake?) later in the day.
